\documentclass{article}
\usepackage{graphicx} % Required for inserting images

\title{Problem Set 4}
\author{adam spohn}
\date{February 2024}

\begin{document}

\maketitle

\section{Data Frame Questions}
what type of object is "mydf(dollar sign)date?
Character.

list the first n rows (i chose 5) of mydf.
 date  description                        lang  category1 category2 granularity
  <chr> <chr>                              <chr> <chr>     <chr>     <chr>      
1 1     Tiberius, under order of Augustus… en    By place  Roman Em… year       
2 1     Gaius Caesar and Lucius Aemilius … en    By place  Roman Em… year       
3 1     Gaius Caesar marries Livilla, dau… en    By place  Roman Em… year       
4 1     Quirinius becomes a chief advisor… en    By place  Roman Em… year       
5 1     Areius Paianeius becomes Archon o… en    By place  Roman Em… year       


What are the class of df and df1?
df is a data frame, df1 is a spark data frame.

Are the column names any different across the two objects? If so, why might that be?
Yes, df1 uses . in between words, df replaces this with an 
undescore so that spark can use it to be compatible with spark sql
because it cannot use dots.

List the first 6 rows of the SepalLength and Species columns of df.

5.1 setosa 
4.9 setosa 
4.7 setosa 
4.6 setosa 
5   setosa 
5.4 setosa 

List the first 6 rows of all columns of df where SepalLength is larger than 5.5.

SepalLength SepalWidth PetalLength PetalWidth Species   
         <dbl>       <dbl>        <dbl>       <dbl> <chr>     
1          5.8         4            1.2         0.2 setosa    
2          5.7         4.4          1.5         0.4 setosa    
3          5.7         3.8          1.7         0.3 setosa    
4          7           3.2          4.7         1.4 versicolor
5          6.4         3.2          4.5         1.5 versicolor
6          6.9         3.1          4.9         1.5 versicolor

\section{Data Sources}
Some data sources I might be interested in scraping from might
include baseball stats sites like baseball savant or baseball 
reference. Baseball savant has a lot of really interesting in
depth stats like diffent pitchers spin rates, that many other 
sites does not have. I also might be intersted in more economy 
based data sites like FRED.

\end{document}
