\documentclass{article}
\usepackage{graphicx} % Required for inserting images

\title{problemset1}
\author{adam spohn}
\date{January 2024}

\begin{document}

\maketitle

    I recently found out about and enrolled into OU's managerial Economics MA/BA program. I have not had much programming experience or emphasis on working with data, but I had thought that if i were to start over that data science might be the route I would go, and entering that program has taken me that route. Along with that has come a new focus on data related jobs after school, though I still feel very new and lacking in experience with data science. For that reason this class fits perfectly with personal interest, and skill expansion in something that am hoping to pursue after graduation. I like with economics how you can see how markets work, but i think the application of data science where you can really find causation is really interesting. 
        I have always liked the movie Moneyball and baseball in general, so once started getting involved in data science working with data for a professional baseball team has sounded really interesting. For this reason doing baseball or sports related research would be interesting to me. In fact I have done a small sports related one for an econometrics class and am working on one now for my economics research class. I have also done a paper over light trucks for a directed readings class that I would be interested to do actually data analytics on. Though really I would just like to find any topic that I feel like could make a deep and meaningful analysis on that can both develop and showcase my data skills to help with any kind of future employment in the data science field. While I think that working for a professional sports team would be really neat, I think that those positions are less common, and maybe  often less enticing than others. Though as of now that is my plan is to look into somewhere that can apply econometric and data science skills. I think this class will be very valuable for supporting that plan. 

        \section {Equations}
        \[a^2 + b^2 = c^2\]

\end{document}
